### 2022-09-07 tex/erosion.tex

\section{Erosion and sedimentation}\label{sec:erosion}
Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
The partial differential equation solved by FastScapeLib is:
\begin{equation}
\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
\end{equation}
where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.

In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
case, continental markers above the top boundary are simply deleted.